---
title: "if-else conditions"
author: "Anita Dhillon"
date: "2023-03-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(magrittr)
library(dplyr)
```

# Logical and conditional statements

These are pieces of code that return ' TRUE' or 'FALSE' values, that is , a logical value 

The common operators of logical statements are
- equality '=='
- inequality '!='
- greater than '>'
- less than '<'
- greater than or equal to '>='
- less than or equal to '<='

The conditional statements allow to test several logical conditions at a time 
The condition operators (or symbols) are 
- AND '&'(inside dplyr function we can represent AND using a ',')
- OR '|'


We also have logical functions that test if something is TRUE or FALSE, for example:
- 'is.na()' : is a function that tests if a value is an 'NA'
    - This function is a part of a whole family of functions that all start with 'is.' : 
    - 'is.vector()'
    - 'is.data.frame()'
    - 'is.factor()'
    
For next class: how to get all functions from a family (method)

- which(): takes logical vectors, it will give you numerical index (position) of all values that are TRUE

```{r}
which(letters == "r")
letters[18]
```


## Exercise 6:Choice Operators

1. Create the following variables
```{r}
 w <- 10.2
 x <- 1.3
 y <- 2.8
 z <- 17.5
 colors <- c("red", "blue", "green")
 masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
 dna1 <- "attattaggaccaca"
 dna2 <- "attattaggaacaca"
```


2. Use them to print whether or not the following statements are TRUE or FALSE.

a) w is greater than 10
```{r}
w > 10
```


b) "green" is in colors
```{r}
("green" == colors)
```


c)x is greater than y
```{r}
x > y
```


d)Each value in masses is greater than 40.
```{r}
is.na(masses > 40)
```

e)2 * x + 0.2 is equal to y

```{r}
2 * x + 0.2 ==  y
```

f)dna1 is the same as dna2

```{r}
(dna1 == dna2)
```

g)dna1 is not the same as dna2
```{r}
dna1 != dna2
```


h)w is greater than x, or y is greater than z
```{r}
w > x | y > z
```

i) x times w is between 13.2 and 13.5
```{r}
x*w < 13.5
x*w > 13.2
# 13.2 < x*w < 13.5 : This is how we would write in on paper
# in R we have to compare things in pairs
x * w < 13.5 & x *w > 13.2

```

j) Each mass in masses is between 30 and 50.

```{r}
masses < 30 & masses < 50 
```


## How to make simple choices w ' if()'

The general structure of an if statement : 

```
if(condition is TRUE) {
   run all lines 
   of code in
   this block
   of code
}
```

If condition is not TRUE, then nothing happens


## Exercise 7 Handing one choice

1. Copy the following code and complete the if statement so that if age_class is equal to “sapling” it sets y <- 10.

```{r}
age_class = "sapling"
 if (age_class == "sapling") {
   y <- 10
 }
y 
```


## Case when we have two options: if-else structure

The general form of this strucuteL 

```
if (condition) {
code that runs if condition IS met
} else {
code that runs if condition is NOT met
}
```

## Exercise 8 : Handling 2 choices

Copy the following code and complete the if statement so that if age_class is equal to “sapling” it sets y <- 10 and if age_class is equal to “seedling” it sets y <- 5.

```{r}
age_class = "seedling"
 if (age_class == "sapling") {
   y <- 10
} else {y <-5 }  
 y
```


## Exercise 9 : Handling two or more choices 

In this case we are using an elseif structure: 

```
if(condition1) {
first block code that executes if condition 1 is met 
} else if (condition2) {
else if (condition 3) {
more code 
} else {
this will cover all the conditions that are not specified before
}
```


1. Copy the following code and complete the if statement so that if age_class is equal to “sapling” it sets y <- 10 and if age_class is equal to “seedling” it sets y <- 5 and if age_class is something else then it sets the value of y <- 0.

```{r}
 age_class = "adult"
 if (age_class == "sapling") {
 (y <- 10) 
 } else if (age_class == "seedling") {
   y<- 5
 } else {
    y <- 0 }
 y
```


## Exercise 10 : value of y by age class


Convert your conditional statement from the last exercise into a function called get_y, that takes age_class as an argument and returns y.

Call the function you just created 5 times, once with each of the following values for age_class, and print the values of y:



## Exercise 12
```{r}
list.files()
if (file.exists("surveys (1).csv")) {
  print("file exists")
} else{
  print("file does not exist")
}
```
?download.file
```{r}
if (file.exists("surveys.csv")) {
  print("file exists")
} else{
  download.file("https://ndownloader.figshare.com/files/2292172", "surveys_test.csv")
  read.csv("surveys_test.csv") %>% 
  head() %>% 
    print()
  
}
```

```{r}
if (file.exists("species1.csv")) {
  print("file exists")
} else{
  download.file("https://ndownloader.figshare.com/files/3299483", "species.csv")
  read.csv("species.csv") %>% 
  head() %>% 
    print()
  
}
```


