---
title: "Joining Data Tables"
author: "Anita Dhillon"
date: "2023-03-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(magrittr)

```

Use the shortcut to add a code block 'ctrl+option + i " on mac ' ctrl + alt i' on windows

Load the three data sets we are going to join , surveys.csv, species, csv, plots.csv 
```{r}
surveys <-read.csv(file = "../data-raw/surveys.csv")
species <-read.csv(file = "../data-raw/species.csv")
plots <- read.csv(file = "../data-raw/plots.csv")
#View(plots)
#View(surveys)
```

### Why do we need to combine or join data tables 


### How do we join data tables in R
there is a group of functions '_join()' that allows us to combine two data tables using values on a shared column 

There has to be a shared column; we need three main arguments to run these functions, two data tbles and one column name

The different functions allows us to combine in different ways .

' inner_join' 

```{r}
inner_join(surveys, species, by = "species_id") %>% 
head()
```

pipe code 

```{r}
surveys |>
inner_join(species, by = "species_id") -> joined_table
```

### How can we explore our combined/joined table 

we want to see the difference between the two input tables and the resulting table
To see the difference in columns we use 'heads()': 

```{r}
head(species)
head(surveys)
head(joined_table)
```

```{r}
str(species)
str(joined_table)
```

What happened with the number of rows in joined_value vs surveys?

It dropped the rows that did not have matching values of the species_id column 

## Exercise 1

```{r}
surveys %>% 
inner_join(plots, by = "plot_id") -> joined_table1
  filter(joined_table1, plot_type == 'Control') %>% 
  head()

```

This returns an error because we tried to join by a coloumn that is not shared by both data tables 

## Automate joining tables/other things with ' intersect'()

```{r}
intersect(surveys$species_id, species$species_id)
```

To find shared columns we use the ' colnames' function 

```{r}
 colnames(surveys)
colnames(species)
intersect(colnames(surveys), colnames(species))
```

Doing it visually 
```{r}
colnames(plots)
colnames(surveys)
```

Automatically with the function 'intersect()'
```{r}
intersect(colnames(surveys),colnames(plots))

```

## Exercise 2

```{r}
inner_join(surveys, plots, by = "plot_id") %>% 
  filter( plot_type == 'Rodent Exclosure') %>% 
  head()

```


### Joining multiple data tables

can we use the '_join()' function on 3 or more tables at the same time (NO)

```{r}
combined <- inner_join(surveys, species, by = "species_id")
combined_final <- inner_join(combined, plots, by = "plot_id")
```

So we can use a Pipe to the join function two or more times (as needed)
```{r}
combined <- surveys |>
inner_join(species, by = "species_id") |>
inner_join(plots, by = "plot_id")
str(combined)
```


### other join Functions 

'left_join()' retains all values from the first table, drops unmatching from second

'right_join()' drops values from the first table and retaining all values from second

'full_join()' keeps all values from both tables



### Exercise 3

```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  filter(taxa == "Rodent") %>% 
  filter(plot_type == "Control"| plot_type == "Long-term Krat Exclosure") %>% 
  filter(!is.na(weight)) %>% 
  select(year, genus, species, weight, plot_type) %>% 
  str()
```



## HOMEWORK (3/13/2023)

```{r}

```







