---
title: "Tree from Open Tree"
author: "Anita Dhillon"
date: "2023-04-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ape) #Analysis,Phylogenetics and Evolution R package
library(phangorn)
library(rotl)
library(magrittr)
library(readr)
```

## Stats 

Insect- 1 million named species
Mammals - 5k named species 
Flowering plants - 200k named species 

### Read a newick tree from Open Tree of Life

The following fucntion can only read trees in newick format. Make sure your file is a newick tree. 

If you got your tree from the phylesystem repo, your tree is going to be in json format and will not be ready 

```{r}
tree <- read.tree(file = "../Data-raw/BunnyTData.tre")
```

Now plot the tree

```{r,fig.height=15 , fig.width=6}
plot(tree, cex= 0.5)
mtext(text = "Oryctolagus cuniculus OpenTree 2017")
```

We can also use the url of a tree to read a tree into R. We will provide the URL in the argument 'file = ' 

In this case, we do not have to keep a copy of the file, we can just download it directly 
The URL is "“http://ape-package.ird.fr/APER/APER2/primfive.tre”



```{r}
small_tree <- read.tree("http://ape-package.ird.fr/APER/APER2/primfive.tre")

```


The R structure of a phylogentic tree

### Getting a tree for your species from the Open Tree of Life 

The open Tree of Life has tools that allow to match the names of a group or species to their unified taxonomy so that we can look for information on their databases. 

If we are doing this in R, we are using the 'tnrs match names ()' from the package 'rtol':

```{r}
tnrs <- tnrs_match_names(names = "amphibians")
class(tnrs)
tnrs
```

What is the main difference between this information that we have in the column 'search_string' and 'unique_name' : 

One is the search string and the other is the scientific taxonmic make from the Open Tree of Life unified Taxonomy (OTT)

We can use the unique number identifier from OTT to get information for that taxon. 

To get a tree we can yse the function 'tol_subtree()'

```{r}
tol_subtree(ott_id = tnrs$ott_id)
```









