---
title: "Joining Vectors"
author: "Anita Dhillon"
date: "2023-03-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(magrittr)
```

## What are vectors ?
 
- They are undimensional matrices
- they can only hold one type of data, either numeric (integer or double), character, or logical (complex numbers)
- it can hold any type of data
- it can only hold different types of data in a certain way: 
  - only columns can have different data types
  - within a coulmn, all rows must have the same data type 
  
  - A data frame can also be identified as a collection of vectors (they can be different or the same type)
  
  

  
```{r}
surveys <- read.csv(file = "./data-raw/surveys.csv")
str(surveys)
```
     
## Creating vectors -review

A character bector with the function 'c'()
```{r}
c("luna", "Avi", "Anita" , "James", "Charles", "Damian", "Davinder") -> our_names
str(our_names)
1:7 # the colon operator creates a vector of numbers 
1:7 -> my_numbers
-100:200
```
     
In how many steps or units does the colon operator increase? In increases in a step of 1

What do we do if I want to create a nummeric sequence that increases in steps different than 1? 
?seq

```{r}
seq(-100,200, by =0.1) ->my_numbers
str(my_numbers)
```
 
 
```{r} 
letters
LETTERS
```
## Creating data frames with vectors  

The most general way to do this is with the function 'data.frame()':

```{r error=TRUE} 
data.frame(names = our_names,numbers = my_numbers)
```

Remeber yectors must have the same length to be part of a data frame:

```{r}
data.frame(names=our_names, numbers = 1:7)
```
     
     
Even though we cant use vectors that are larger to create a data frame, we can use vector that are shorter, and R will recycle the values: 

```{r}
data.frame(names = our_names,numbers =1)
```
     
To recycle the values of a numeric vector length 2, we have to repeat the vector of names two times so it is a multiple of 2: 

```{r}
data.frame(names = rep(our_names, 2), numbers = c(2, 5.5))
```



## Exercise 6

```{r}
length <- c(2.2, 2.1, 2.7, 3.0, 3.1, 2.5, 1.9, 1.1, 3.5, 2.9)
width <- c(1.3, 2.2, 1.5, 4.5, 3.1, NA, 1.8, 0.5, 2.0, 2.7)
height <- c(9.6, 7.6, 2.2, 1.5, 4.0, 3.0, 4.5, 2.3, 7.5, 3.2)
```     
     
  
```{r} 
data.frame(length,width,height) -> table
table$genus <- "Taxus"
table$species <- "baccata"
str(table)
table
```
  
  
###Extracting/accessing values from vectors and data frames

```{r}
surveys <- read.csv(file = "./data-raw/surveys.csv")
str(surveys)
```  


```{r}
surveys$record_id -> record_id
```  
  
  
Another way is with the square brackets '[]'

```{r}
surveys[1:10,"hindfoot_length"]
```

If i want all the values from the rows of column hindfoot length

```{r}
surveys[ , "hindfoot_length" ] -> hindfoot_length 
head(hindfoot_length)

```


Another way is using double square brackets 

```{r}
surveys[["record_id"]] %>% 
  head()
```

### Exercise 7

```{r}
surveys$weight -> surveys_weight
surveys["month"] -> surveys_month
head(surveys_weight)
```



```{r}
surveys[,"hindfoot_length"] %>% 
head()
mean(surveys$hindfoot_length,na.rm = TRUE) %>% 
head()
```





